---
title: "ENVS 193DS Final"
author: "Autrin Peralta"
format: html
Date: Winter 2026
editor: visual
---

<https://github.com/autrinp/Winter-2026_Final>

```{r}
#| label: Loading in packages
#| message: false
library(tidyverse)
library(janitor)
library(readr)
kelp <- read_csv("../Data/Annual_Kelp_All_Years_20250903.csv")
```

# Problem 1

a\.

-   In part 1 they used simple linear regression technique as there was a predictor variable(precipitation) and a response variable(flooded wetland area. Also it doesn't predict a relationship showing there's a relationship between the two variables

-   In part 2 they used a non-parametric test as it looks to measure the "median" flooded wetland area. When measuring 5 predictor variables, they used a Kruskal-Wallis rank sum test.

b\.

-   Measuring the median values, a box plot would be the most appropriate for part 2. The x-axis would have the water year classifications and the y-axis would have the response variable(flooded wetland area). Each group would be represented by a box with a middle line(the median), whiskers to emphasize the range of the data, and the IQR which shows the spread of the data.

c\.

-   Part 1: Precipitation doesn't predict flooded wetland area in the San Joaquin River Delta (linear regression, F(df of model) = f statistic value, R\^2 = determination, p = 0.11, 𝛂 = significance level).

-   Part 2: We found no difference in median flooded wetland area across different water year classification (wet, above normal, below normal, dry, critical drought)(Kruskal_Wallis test, X²(distribution)(df) = test statistic value, p = 0.12, 𝛂 = significance level)

d\.

-   River/tide height(continuous variable) could affect flooded wetland area. Higher water levels means nothing is stopping the water from flooding the nearby areas. Management factors may play a role in this as they may control how much water is allowed in during high tide seasons.

# Problem 2

a\.

```{r}
#| label: Cleaning and Summarizing data
#| warning: false
kelp_clean <- kelp |> #creating new data set from "kelp"
  clean_names() |> #clean names to be all lower case with _
  filter(site%in% c("IVEE", "NAPL", "CARP" )) |> #filter to only include "IVEE", "NAPL", and "CARP"
  mutate(site_name = case_when( #mutate the existing site_name column
    site == "NAPL" ~ "Naples", #Fill in "Naples" when "NAPL" appears
  site == "IVEE" ~ "Isla Vista", #Fill in "Isla Vista" when IVEE appears
  site == "CARP" ~ "Carpinteria" #Fill in "Carpenteria" when "CARP" appears
  )) |> 
  filter(!(date == "2000-09-22" & transect == "6" & quad == "40")) |>  #negating this row of data 
   group_by(site_name, year) |> #group by site name and year
  summarize(mean_fronds = mean(fronds)) |> # calculate the mean
  ungroup() |> #ungroup the data
 mutate(site_name = as_factor(site_name), #makes the categorical variable into a factor so it can be used in models
       site_name = fct_relevel(site_name, #relevel gives order 
                               "Naples", "Isla Vista", "Carpinteria"))
  

  
```

```{r}
#| label: Slicing data
#| message: false
kelp_clean |> # pulling from "kelp_clean" data
  slice_sample(n = 5) #pulling 5 random data points
```

```{r}
kelp_clean |> #pulling from data
  str() #showing structureed data
```
