---
title: "ENVS 193DS Final"
author: "Autrin Peralta"
format: html
Date: Winter 2026
editor: visual
---

<https://github.com/autrinp/Winter-2026_Final>

```{r}
#| label: Loading in packages
#| message: false
library(tidyverse)
library(janitor)
library(readr)
library(DHARMa)
library(ggeffects)
kelp <- read_csv("../Data/Annual_Kelp_All_Years_20250903.csv")
mopl <- read_csv("../Data/MOPL_nest-site_selection_Duchardt_et_al._2019.csv")
```

# Problem 1

a\.

-   In part 1 they used simple linear regression technique as there was a predictor variable(precipitation) and a response variable(flooded wetland area. Also it doesn't predict a relationship showing there's a relationship between the two variables

-   In part 2 they used a non-parametric test as it looks to measure the "median" flooded wetland area. When measuring 5 predictor variables, they used a Kruskal-Wallis rank sum test.

b\.

-   Measuring the median values, a box plot would be the most appropriate for part 2. The x-axis would have the water year classifications and the y-axis would have the response variable(flooded wetland area). Each group would be represented by a box with a middle line(the median), whiskers to emphasize the range of the data, and the IQR which shows the spread of the data.

c\.

-   Part 1: Precipitation doesn't predict flooded wetland area in the San Joaquin River Delta (linear regression, F(df of model) = f statistic value, R\^2 = determination, p = 0.11, 𝛂 = significance level).

-   Part 2: We found no difference in median flooded wetland area across different water year classification (wet, above normal, below normal, dry, critical drought)(Kruskal_Wallis test, X²(distribution)(df) = test statistic value, p = 0.12, 𝛂 = significance level)

d\.

-   River/tide height(continuous variable) could affect flooded wetland area. Higher water levels means nothing is stopping the water from flooding the nearby areas. Management factors may play a role in this as they may control how much water is allowed in during high tide seasons.

# Problem 2

a\.

```{r}
#| label: Cleaning and Summarizing data
#| warning: false
kelp_clean <- kelp |> #creating new data set from "kelp"
  clean_names() |> #clean names to be all lower case with _
  filter(site%in% c("IVEE", "NAPL", "CARP" )) |> #filter to only include "IVEE", "NAPL", and "CARP"
  mutate(site_name = case_when( #mutate the existing site_name column
    site == "NAPL" ~ "Naples", #Fill in "Naples" when "NAPL" appears
  site == "IVEE" ~ "Isla Vista", #Fill in "Isla Vista" when IVEE appears
  site == "CARP" ~ "Carpinteria" #Fill in "Carpenteria" when "CARP" appears
  )) |> 
  filter(!(date == "2000-09-22" & transect == "6" & quad == "40")) |>  #negating this row of data 
   group_by(site_name, year) |> #group by site name and year
  summarize(mean_fronds = mean(fronds)) |> # calculate the mean
  ungroup() |> #ungroup the data
 mutate(site_name = as_factor(site_name), #makes the categorical variable into a factor so it can be used in models
       site_name = fct_relevel(site_name, #relevel gives order 
                               "Naples", "Isla Vista", "Carpinteria"))
  

  
```

```{r}
#| label: Slicing data
#| message: false
kelp_clean |> # pulling from "kelp_clean" data
  slice_sample(n = 5) #pulling 5 random data points
```

```{r}
kelp_clean |> #pulling from data
  str() #showing structureed data
```

b\.

This data set was deleted because it was -99999 which indicates missing or not needed data points. Found in "methods" tab of data page.

c\.

```{r}
ggplot(data = kelp_clean, #creating plot from data set "kelp_clean"
       aes(x = year, #x-axis
           y = mean_fronds, #y-axis
           color = site_name)) + #color by site na,e
  geom_point() + #add points for data
  geom_line() + #connect poionts with a line
  scale_color_manual(
    values = c("Naples" = "turquoise", #color Naples points
               "Isla Vista" = "tan", #color Isla Vista points
               "Carpinteria" = "darkblue" #color carpenteria points
               )) +
  labs(x = "Year", #rename x-axis
       y = "Mean kelpfFronds per year", #rename y-axis
       title = "Sites vary in mean kelp fronds per year") + #add title
  theme_minimal() + #change background to white
  theme(
  panel.border = element_blank(), #eliminate border
  plot.title = element_text(hjust = 0.5), #center title
  text = element_text(family = "Times New Roman")) #change font to times new roman
  
  
  
```

d\.

-   Isla vista looks like it has the the the highest and lowest peaks and most variability between each peak

-   Carpinteria has the flattest set of data showing less variability.

# Problem 3

a\. The 1's and 0's represent whether a nest site is being used by the Mountain Plover or not. 1's indicate a used site and 0's indicate an available site.

b\. Visual obstruction(continuous) - "selected model indicated that plovers avoided nesting in areas with a maximum vegetation height \>11 cm, corresponding with an average visual obstruction of \>5 cm."(Discussion, paragraph 4). As height of the vegetation increases(increasing visual obstruction) thsi decreases the chance of a plover to nest there as they prefer open areas.

Distance to prairie dog colony(continuous) - "this is the first direct evidence that colony size may negatively influence plover density"(Discussion, paragraph 2). As distance to colony edge increases(colony size), the probability of Mountain Plover nest site use would decrease.

c\.

|  |  |  |  |  |  |  |
|--------------|------|---------------|------|--------------|------|-------------|
| **Model Name** |  | **Visual Onbstruction** |  | **Distance to praire colony** |  | **Model Description** |
| mod_null |  |  |  |  |  | No predictor(fmodel) |
| Mod_sat |  | X |  | X |  | All predictors(model |
| Mod_Edgedist |  |  |  | X |  | Distance only |
| Mod_vo |  | X |  |  |  | Visual Only |

d\.

```{r}
mopl_clean <- mopl |> # creating a new data setmopl_clean
  clean_names() |> # cleaning the names
  mutate(used_bin = ifelse(id == "used",1, 0)) |> # making a new column to show 1 if nest is used and 0 if nest isnt used
  select(used_bin, edgedist, vo_nest) #selecting bins to use
```

```{r}
mod_null <- glm(used_bin ~ 1, #making new object "mod_null", has no predictors
             data = mopl_clean, # data
             family = binomial) #error distribution

```

```{r}
mod_sat <- glm(used_bin ~ edgedist + vo_nest, # making new object mod_sat, both edgedist and vo as predictors
               data = mopl_clean, # data
              family = binomial) #error distribution
```

```{r}
mod_edgdist <- glm(used_bin ~ edgedist, # making new object mod_edge, edgedist as predictor
                data = mopl_clean,# data
                family = binomial) #error distribution
```

```{r}
mod_vo <- glm(used_bin ~ vo_nest, # making new object mod_vo, vo as predictor
              data = mopl_clean,# data
             family = binomial) #error distribution
```

e\.

```{r}
AIC(mod_sat,
    mod_null,
    mod_vo,
    mod_edgdist) # calculate the AIC for each model
```

The best model as determined by Akaike’s Information Criterion (AIC) is mod_vo for obstruction of visualization(AIC = 363.4420).

f\.

```{r}
plot(simulateResiduals(mod_vo)) # diagnostic plot
```

g\.

```{r}
mod_preds <- ggpredict(mod_vo,
                       terms = "vo_nest[all]") # predicted probabilites
ggplot(mopl_clean, # using mopl_clean data frame
       aes(x = vo_nest, # set x-asix
           y = used_bin, #set y-axis
           color = "Observed")) + # colored data points
  geom_point(aes(),
             size = 3, 
             alpha = 0.4) + # coloring data points)
  geom_ribbon(data = mod_preds, # using mod_preds for ribbon data
              aes(x = x, # same x-axis
                  y = predicted, 
                  ymin = conf.low,
                  ymax = conf.high),
              fill = "purple", # Setting color 
              alpha = 0.4) + 
  geom_line(data = mod_preds, # using mod_preds data for line
            aes(x = x, #same x-axis
                y = predicted,
                color = "Model"), # coloring the line and setting width
            linewidth = 1) +
  labs(x = "Nest visual obstruction (cm)",  # x-axis name
  y = "Probability of nest occupancy") + #y-axis name
  scale_y_continuous(limits = c(0, 1), # setting y-axis scale
                     breaks = c(0, 1)) + 
  scale_color_manual(values = c("Observed" = "tan", # manually coiloring the line and data points
                                "Model" = "darkgreen")) +
  theme_minimal() + # changing background white
theme(panel.grid = element_blank())  # removing gridlines
```

h\.

# Problem 4

a\.

-   The data from HW 3 contains a ton more data points, HW 2 only has 4 points of datawhile HW 3 has 30 observations. HW 3 has colored points for the predictor variables( green = no flip flops, red = flip flops) and is a lot cleaner and organized.

-   Both are jittered plots with the response varibale on the y axis(biking time and sleeping time).
